### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f7b24e3dc31af379be75123d05d7c99049d263fe 🚀

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South
America has the most evolutionary distinct hosts \emph{and} viruses,
followed by secondary hotspots in southeast Asia and the Rift Valley
region have mostly distinct viruses. Some degree of sampling bias may
contribute to these patterns: for example, South-America is one of the
places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{2,13,15} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{16}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness.

These patterns recapitulate the evolutionary history of both the order
Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe bats
(Rhinolophidae) are both the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}) and the presumed ancestral host of
\emph{Betacoronavirus}.\textsuperscript{17} The hotspot of host richness
and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---is almost entirely driven by adaptive radiation in this
clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
viral distinctiveness is the result of isolation by vicariance for both
bats and their viruses. Out of the four main groups of
betacoronaviruses, only the subgenus \emph{Merbecovirus} has been found
in animals in the Americas---an introduction that is generally presumed
to be ancient. While comparatively understudied, New World
merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
groups are endemic to the Neotropics, while the latter two and
particularly the phyllostomids are responsible for the hotspot of bat
diversity in the Amazon. Together, these groups play host to a distinct
regime of bat-betacoronavirus coevolution in the Americas.

>>>>

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South
America has the most evolutionary distinct hosts \emph{and} viruses,
followed by secondary hotspots in southeast Asia and the Rift Valley
region have mostly distinct viruses. Some degree of sampling bias may
contribute to these patterns: for example, South-America is one of the
places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{2,13,15} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{16}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

On closer inspection, these patterns recapitulate the evolutionary
history of both the order Chiroptera and the genus
\emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}) and
the presumed ancestral host of
\emph{Betacoronavirus}.\textsuperscript{17} The hotspot of host richness
and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---is almost entirely driven by adaptive radiation in this
clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
viral distinctiveness is the result of isolation by vicariance for both
bats and their viruses. Out of the four main groups of
betacoronaviruses, only the subgenus \emph{Merbecovirus} has been found
in animals in the Americas---an introduction that is generally presumed
to be ancient. While comparatively understudied, New World
merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
groups are endemic to the Neotropics, while the latter two and
particularly the phyllostomids are responsible for the hotspot of bat
diversity in the Amazon. Together, these groups play host to a distinct
regime of bat-betacoronavirus coevolution in the Americas.
